---
title: "NEON Project"
author: "Sarah Munro"
output: html_notebook
---


## Install packages
```{r}
library(tidyverse)
library(neonUtilities)
library(lubridate)
```

## Load Data Periphyton, seston, phytoplankton chemical properties
```{r}
#PeriphytonSestonPhytoplanton <- loadByProduct(dpID = "DP1.20163.001", package = "basic", site = c("ARIK"))
```

```{r}
#save(PeriphytonSestonPhytoplanton, file= "data/PeriphytonSestonPhytoplanton.Rdata")
load("data/PeriphytonSestonPhytoplanton.Rdata")


```

## Load PAR
```{r}
#PARwatersurface <- loadByProduct(dpID = "DP1.20042.001", package = "basic", site = c("ARIK"))
```

```{r}
# save(PARwatersurface, file= "data/PARwatersurface.Rdata")
load("data/PARwatersurface.Rdata")


```

#Look at table
```{r}
ARIKBioMass <- PeriphytonSestonPhytoplanton$alg_algaeExternalLabDataPerSample
ARIKHabitat <- PeriphytonSestonPhytoplanton$alg_fieldData
```

#Remove columns with NA
```{r}
ARIKBioMass <- filter(ARIKBioMass,!is.na(ARIKBioMass$analyteConcentration))
ARIKBioMass <- filter(ARIKBioMass,!is.na(ARIKBioMass$sampleCode))
ARIKBioMass <- filter(ARIKBioMass,!is.na(ARIKBioMass$acidTreatment))
```
#Remove unwanted columns
```{r}
ARIKBioMass <- ARIKBioMass %>%
  select(uid, siteID, collectDate, sampleID, sampleCondition, analysisDate, analyte, analyteConcentration, plantAlgaeLabUnits, testMethod)
ARIKHabitat <- ARIKHabitat %>%
  select(uid, domainID, siteID, aquaticSiteType,decimalLatitude, decimalLongitude, collectDate, parentSampleID, habitatType, algalSampleType)
```

## Make decimal date and put into table
```{r}
ARIKBioMass <-
  ARIKBioMass %>%
  mutate(decimaldate = decimal_date(ARIKBioMass$collectDate))
ARIKHabitat <-
  ARIKHabitat %>%
  mutate(decimaldate = decimal_date(ARIKHabitat$collectDate))
```

## Put carbon into one table
```{r}
carbon <-ARIKBioMass %>%
  filter(analyte == c("carbon"))
```

# Plotting carbon 
```{r}
ggplot(data =carbon) +
  geom_line(aes(x = decimaldate, y = analyteConcentration))
```
# PLOTTING CARBON FROM 7/18 - 3/21

# AVG carbon
```{r}
carbon$Date <- as.Date(carbon$collectDate)

AVG_Carbon <- carbon %>%
  group_by(Date) %>%
  summarise(analyteConcentration = mean(analyteConcentration))
```

# GRAPH 2 AVG CARBON JULY 2018- MARCH 2021
```{r}
AVG_Carbon %>%
  ggplot(aes(x = Date, y= analyteConcentration)) +
         geom_col()
```


# PAR


## Load Data Photosynthetically active radiation (PAR)
```{r}
# PAR <- loadByProduct(dpID = "DP1.00024.001", package = "basic", site = c("ARIK"))
# as of12/6 changed to dif dataset, loaded above
```
#Save PAR data
```{r}
#save(PAR, file= "data/Photosynactrad.Rdata")
load("data/PARwatersurface.Rdata")

```
# Look at table
```{r}
ParWtr5 <- PARwatersurface$PARWS_5min
Parwtr30 <- PARwatersurface$PARWS_30min
Parwtr30
```
##Remove rows with NA 
```{r}
Par_wtr30<- filter(Parwtr30, !is.na(Parwtr30$PARMean))
Par_wtr30

```

# remove unnecessary cols
```{r}
Par_wtr30 <- Par_wtr30 %>%
  select(startDateTime, PARMean, PARMinimum, PARMaximum, PARStdErMean)
Par_wtr30
```

# add a column for decimal data
```{r}
Par_wtr30 <-
  Par_wtr30 %>%
  mutate(decimaldate = decimal_date(Par_wtr30$startDateTime))
Par_wtr30

```


# filter from july 2018- march 2021 to overlap w biomass
```{r}
Par_wtr30 <- Par_wtr30 %>%
  filter(between(decimaldate, 2018.496, 2021.162 ))

Par30_coupledays <- Par30_coupledays %>%
  filter(between(decimaldate, 2018.496, 2021.162 ))

```



## PLOTTING MEAN OVER TIME OF 30 MIN INTERVAL
```{r}
Par_wtr30 %>%
    ggplot(aes(x = decimaldate,y = PARMean)) +
  geom_line()

```


## FILTER PAR IN 30MIN FOR A COUPLE DAYS
```{r}
Par30_coupledays <- Par30_coupledays%>%
  filter(between(decimaldate, 2018.496, 2018.504))
```

#PLOT couple days
```{r}

Par30_coupledays %>%
    ggplot(aes(x = decimaldate,y = PARMean)) +
  geom_line()


```

# adding column #AVG PAR GRAPH 3 OVERLAP W CARBON JULY 2018-MARCH 2021
```{r}
Par_wtr30$Date <- as.Date(Par_wtr30$startDateTime)

AVG_Par <- Par_wtr30 %>%
  group_by(Date) %>%
  summarise(meanPar = mean(PARMean))


AVG_Par %>%
  ggplot(aes(x = Date, y = meanPar, color = "blue")) +
  geom_line()
```

# PAR & PRECIP filter 5/14-5/19
```{r}
Parwtr30 <-
  Parwtr30 %>%
  mutate(decimaldate = decimal_date(Parwtr30$startDateTime))
Parwtr30


Parwtr30 <- Parwtr30 %>%
  filter(between(decimaldate,2016.366, 2016.382 ))

PRECIP_30$Date <- as.Date(PRECIP_30$startDateTime)

SumOfPrecip <- PRECIP_30 %>%
  group_by(Date) %>%
  summarise(priPrecipBulk = sum(priPrecipBulk))

Parwtr30$Date <- as.Date(Parwtr30$startDateTime)

AVGPatwtr30_precip <- Parwtr30 %>%
 group_by(Date) %>%
  summarise(meanPar = mean(PARMean,na.rm = TRUE))

```
 
 # PLOT PRECIP 5/14-5/19
```{r}
SumOfPrecip %>%
  ggplot(aes(x = Date, y = priPrecipBulk, color = "blue")) +
  geom_line()
```
 # PLOT AVG PAR 5/14-5/19
```{r}
AVGPatwtr30_precip %>%
  ggplot(aes(x = Date, y = meanPar, color = "blue")) +
  geom_line() +
  scale_x_date(limits = as.Date(c("2016-05-14", "2016-05-19")))
```
 
```
 


``

# limiting rows
```{r}
df2 <- ARIKPar1[-(161:3333679),]

ARIK_PAR1 <- df2

df2 <- ARIKPar30[-(161:111172),]

ARIK_PAR30 <- df2

```


### Load Data Precipitation
```{r}
Precipitation <- loadByProduct(dpID = "DP1.00006.001", package = "basic", site = c("ARIK"))
```
## Save data
```{r}
#save(Precipitation, file= "data/PrecipitationData.Rdata")
load("data/PrecipitationData.Rdata")
```

#Look at table 
```{r}
PRECIP30 <- Precipitation$PRIPRE_30min
PRECIP5 <- Precipitation$PRIPRE_5min
```

#REMOVE NA
```{r}
PRECIP30<- filter(PRECIP30, !is.na(PRECIP30$priPrecipBulk))

PRECIP5<- filter(PRECIP5, !is.na(PRECIP5$priPrecipBulk))
```

#removing 
```{r}
df2 <- PRECIP5[-(161:585240),]

PRECIP_5 <-df2

df2 <- PRECIP30[-(161:97575),]

PRECIP_30 <- df2

```
 
```{r}

```

```
 

